---
title: "In-Class Exercise - "

author: "Pierre HAAS"

date: "February 13, 2023"
date-modified: "`r Sys.Date()`"

execute:
  eval: true
  echo: true
  warning: false
editor: visual

number-sections: true
---

# Getting Started

## Installing and Loading Packages

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse)
```

## Import data sets

### Aspatial data

```{r}
hunan2012 = read_csv("data/aspatial/Hunan_2012.csv")
head(hunan2012)
```

### Geospatial data

```{r}
hunanGeo = st_read(dsn = "data/geospatial",
                   layer = "Hunan")
glimpse(hunanGeo)
```

## Combining data frames

```{r}
hunan_gdppc = left_join(hunanGeo, hunan2012) %>%
  select(1:4, 7, 15)
```

Note that we keep columns *NAME_3* and *County* to double check that the left join was done correctly and that spelling is the same across the two columns.

## Plotting a Choropleth map

```{r}
tm_shape(hunan_gdppc)+
  tm_fill("GDPPC",
          style = "quantile",
          palette = "Blues",
          title = "GDP / cap")+
  tm_layout(main.title = "Distribution of GDP per capita by ",
          main.title.position = "center",
          main.title.size = 1.2,
          legend.height = 0.45,
          legend.width = 0.35,
          frame = TRUE)+
  tm_borders(alpha = 0.5)+
  tm_compass(type = "8star", size = 2)+
  tm_scale_bar()+
  tm_grid(alpha = 0.2)
  
```

## Contiguity Neighbors Analysis

Using the Queen method

```{r}
cn_queen = hunan_gdppc %>%
  mutate(nb = st_contiguity(geometry),
         .before = 1)
```

Using the Rooks method

```{r}
cn_rooks = hunan_gdppc %>%
  mutate(nb = st_contiguity(geometry, queen = FALSE),
         .before = 1)
```

## Computing Contiguity Weights

Using the Queen method

```{r}
wm_q = hunan_gdppc %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb),
         .before = 1)
```

Using the Rooks method

```{r}
wm_r = hunan_gdppc %>%
  mutate(nb = st_contiguity(geometry),
         queen = FALSE,
         wt = st_weights(nb),
         .before = 1)
```
