---
title: "Hands-on Exercise - Session 1: Geospatial Data Wrangling with R"
execute: 
  warning: false
date: "`r Sys.Date()`"
---

# Getting Started

## Installing and Loading R Packages

In this section, I will install and load the required R packages, being tidyverse and sf libraries/packages.

```{r}
pacman::p_load(tidyverse, sf)
```

# Importing Geospatial Data in shp format

## Import Data

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
cpg <- st_read(dsn = "data/geospatial", 
                layer = "CyclingPathGazette")
```

## Checking Content of Data Frame Simple Feature

```{r}
st_geometry(mpsz)
```

```{r}
glimpse(mpsz)
```

## Plotting Geospatial Data - Polygon Feature

```{r}
plot(mpsz)
```

## Plotting Geospatial Data - Polyline Feature

```{r}
plot(cpg)
```

# Importing GIS data in kml format

```{r}
psl <- st_read(dsn = "data/geospatial/preschools-location.kml")
```

## Plotting Geospatial Data - Multipoint

```{r}
plot(psl)
```
