---
title: "Hands-on Exercise - Choropleth Mapping with R"
subtitle: "Session 3"
execute: 
  eval: true
  echo: true
  warning: false
---

# 1 Getting Started

## 1.1 Installing and Loading R Packages

In this section, I will install and load the required R packages, tidyverse, sf, and tmap packages. Here we use the pacman library to install and load the packages at once.

```{r}
pacman::p_load(tidyverse, sf, tmap, stringr)
```

# 2 Importing Data into R

## 2.1 Importing Geospatial Data

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## 2.2 Importing Attribute Data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv") %>%
  filter(Time==2020)
```

# 3 Data Preparation

In this section, we will start off by preparing a data table with year 2020 values. The data table will include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY. Please find below a description of the mentioned variables:

-   YOUNG: age group 0 to 4 until age group 20 to 24

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64

-   AGED: age group 65 and above

-   TOTAL: all age group

-   DEPENDENCY: the ratio between young and aged against economy active group

## 3.1 Data Wrangling

```{r}
popdata2020 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

## 3.2 Joining the Geospatial and Attribute Data

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

# 4 Choropleth Mapping for Geospatial Data using *tmap*

## 4.1 Quickly plot a choropleth map with *qtm()* function

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

## 4.2 Creating a choropleth map using *tmap's* elements

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

### 4.2.1 Drawing a base map

```{r}
tm_shape(mpsz_pop2020) + 
  tm_polygons()
```

### 4.2.2 Drawing a choropleth map using *tm_polygons()*

```{r}
tm_shape(mpsz_pop2020) + 
  tm_polygons("DEPENDENCY")
```

### 4.2.3 Drawing a choropleth map using *tm_fill()* and *tm_border()*

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1, alpha = 1)
```

N.B. Some important things to note about the *tm_fill()* function:

The alpha argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).

Beside alpha argument, there are three other arguments for tm_borders(), they are:

-   *col* = border colour,

-   *lwd* = border line width. The default is 1, and

-   *lty* = border line type. The default is "solid".

## 4.3 Data classification methods of *tmap*

### 4.3.1 Plotting choropleth maps using built-in classification methods

The code chunk below shows a quantile data classification that used 5 classes. In order words, we specify using the *jenks* method with 5 classes.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

In the code chunk below, *equal* data classification method is used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

#### DIY

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Show the code"
#| label: fig-mpg
#| fig-cap: "Distribution of Dependency Ratio by Planning Subzone."
#| fig-subcap:
#|   - "Quantile classification"
#|   - "Standard deviation classification"
#|   - "Pretty classification"
#|   - "Kmeans classification"
#| layout-ncol: 2
#| column: page

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5)

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "sd") +
  tm_borders(alpha = 0.5)

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "pretty") +
  tm_borders(alpha = 0.5)

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans") +
  tm_borders(alpha = 0.5)
```
